Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-29 article.tex
add values for phi1 and phi2
Section A.3 rewritten

This heuristic is based on the following observation. The rise of fluorescence
of a neuron indicates its activation. If another neuron is activated after a
slight delay, this could be a consequence of the activation of the first neuron
and therefore indicate a directed link in the connectome from the first to the
second neuron. Given this analysis, the idea of our heuristic is to count the
number of such events occurring between neurons $i$ and $j$ for consecutive
time steps. Mathematically, we compute for each pair $(i,j)$:
\begin{align}
s_{i,j} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
where $\phi_1$ and $\phi_2$ are parameters of the heuristic.

Two issues remain with this heuristic. First, it does not allow to detect a
directed relationship if both neurons are activated simultaneously at the same
time step, i.e., when the delay between the two activations is below the time
resolution. Second, one needs to compare $s_{j,i}$ and $s_{i,j}$ to decide on
edge orientation. Indeed, edge orientation can only be predicted with some
confidence if either $s_{i,j}$ is greater than $s_{j,i}$ ($i\rightarrow j$) or
$s_{i,j}$ is lower than $s_{j,i}$ ($j\rightarrow i$). To address these issues,
we computed a differential pairwise coefficient $z_{i,j}$ defined as follows:
\begin{align}
z_{i,j} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.

\textcolor{red}{PG: je ne comprends pas ce qui est fait exactement. Mettre les
  maths? Valeurs de $\phi_1$ et $\phi_2$ ?} At the light of the poor results of this heuristic taken on its own as
\textit{directivity}
(on Table \ref{tab:directivity} in Section~\ref{sapp:results}), we use it to asymmetrize the degree of association
matrix using partial correlation statistics. We expect that when the heuristic
is highly confident (i.e., $z_{i,j} \gg 0 \gg z_{j,i}$) the degree of
association of edge $(i,j)$ is increased while the degree of association
corresponding to the edge $(j,i)$ is reduced. Even if this adjustment is
subtle, it might have a slight positive effect on the ROC AUC.

>>>>

This heuristic is based on the following observation. The rise of fluorescence
of a neuron indicates its activation. If another neuron is activated after a
slight delay, this could be a consequence of the activation of the first neuron
and therefore indicate a directed link in the connectome from the first to the
second neuron. Given this analysis, the idea of our heuristic is to count the
number of such events occurring between neurons $i$ and $j$ for consecutive
time steps. Mathematically, we compute for each pair $(i,j)$:
\begin{align}
s_{i,j} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
where $\phi_1$ and $\phi_2$ are parameters of the heuristic (with $\phi_1 = 0.2$ and $\phi_2 = 0.5$ in our experiments).

Two issues remain with this heuristic. First, it does not allow to detect a
directed relationship if both neurons are activated simultaneously at the same
time step, i.e., when the delay between the two activations is below the time
resolution. Second, one needs to compare $s_{j,i}$ and $s_{i,j}$ to decide on
edge orientation. Indeed, edge orientation can only be predicted with some
confidence if either $s_{i,j}$ is greater than $s_{j,i}$ ($i\rightarrow j$) or
$s_{i,j}$ is lower than $s_{j,i}$ ($j\rightarrow i$). To address these issues,
we computed a differential pairwise coefficient $z_{i,j}$ defined as follows:
\begin{align}
z_{i,j} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters. The coefficient $z_{i,j}$ estimates the confidence in an edge $i\rightarrow j$. Note that if $z_{i,j}$ is positive, then $z_{j,i}$ will be negative.

This heuristic does not give good results on its own (see Table \ref{tab:directivity} in Section~\ref{sapp:results}). However, it can be used to asymmetrize the degree of association (symmetric) matrix given by partial correlation statistics. Concretely, a new matrix is computed as the linear combination of the partial correlation matrix and the heuristic matrix. 
Mathematically, we have
\begin{align}
q_{i,j} = weight_p * p_{i,j} + weight_z * z_{i,j}
\end{align}
where $weight_p$ and $weight_z$ are parameters of the method (with $weight_p = 0.997$ and $weight_z = 0.003$ in our experiments).

Considering the small weight associated to this heuristic, the effect is subtle. It slightly increases the confidence in one edge direction and decreases the confidence in the opposite edge direction while not affecting strongly the undirected topology of the network. This process slightly improves the performance of our method (see Table \ref{tab:directivity} in Section~\ref{sapp:results}).
